SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-05 back.tex
Adding Perc image, motivation section

Compared to some of the other state of the art prefetchers, SPP is
less aggressive.  In a single-core environment, there is no resource
contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
As we increase the core
count, we observe that SPP starts outperforming rest of the prefetchers.  This
can be attributed to the fact that each prefetch suggested by SPP is a
carefully calculated and prevents cache pollution.  \textit{<Figure
showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.

The above analysis shows that with a more careful filtering mechanism, any
prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
independent filter.  To test the hypothesis, we tuned down SPP to the
minimum possible threshold, with the effect of increasing prefetch suggestions
made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
Obviously this came at a cost of increased DRAM traffic and cache pollution.

Moreover, the on-line confidence mechanism introduced in SPP was very
rudimentary.  It was based on taking a ratio C\textsubscript{d} =
C\textsubscript{delta} / C\textsubscript{sig} as explained previously. This
confidence was used to make the decision of whether to prefetch or not to
prefetch; and which level to prefetch.  While this approximation was shown to
work in the original implementation, we believe that a better form of
generalized on-line decision making was possible.  Hence, it was necessary to
build a robust and adaptable learning mechanism to accept / reject the
prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level
Cache).

@@

Perceptrons in microarchitecture were first introduced by Jimenez and Lin ~\cite{Perc_Branch} for branch prediction. In this work, a perceptron weights vector is selected by a hash of the branch address and its dot product with an input vector of branch history outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is predicted not taken. The weights are updated with the perceptron update rule: if the prediction is incorrect, or the dot product does not exceed some threshold magnitude, then the weights are incremented or decremented based on their correlation to the corresponding history bits. Perceptron learning yields to superior accuracy, which results in increased opportunity for optimizations.

@@

In a perceptron-based mechanism, a uniform
perceptron update principle is followed.  Perceptrons are a simple model of neurons
in neural networks \cite{perceptrons1,perceptrons2} modeled by vectors of signed weights learned through online training. The output of a perceptron is the dot product of the weights and a vector of inputs. In this work, we make use of the perceptron learning algorithm. There are two components to the abstract perceptron learning algorithm:

>>>>

%% Compared to some of the other state of the art prefetchers, SPP is
%% less aggressive.  In a single-core environment, there is no resource
%% contention among the different cores, hence aggressive prefetching is bound to prove beneficial. 
%% As we increase the core
%% count, we observe that SPP starts outperforming rest of the prefetchers.  This
%% can be attributed to the fact that each prefetch suggested by SPP is a
%% carefully calculated and prevents cache pollution.  \textit{<Figure
%% showing the variation of SPP vs BOP wrt core count>} For 4-core applications,
%% SPP suggests XX\% fewer prefetches than Best Offset Prefetcher ~\cite{bop} and yet leads to higher IPC.
%% 
%% The above analysis shows that with a more careful filtering mechanism, any
%% prefetcher like SPP can be tuned to become much more aggressive, yielding higher coverage.  The responsibility of maintaining high accuracy now falls on the
%% independent filter.  To test the hypothesis, we tuned down SPP to the
%% minimum possible threshold, with the effect of increasing prefetch suggestions
%% made by SPP by XX\%.  \textit{<Comparison of SPP-Unleased with SPP / BOP>}
%% Obviously this came at a cost of increased DRAM traffic and cache pollution.

% EB: This remains one of the untouched subsections of the paper

As was noted in Figure~\ref{Fig:Motivation}, an aggressive lookahead 
prefetching if done without any accuracy check, can harm the performance 
of the system. In this given case, the effective performance gained
by the original prefetcher is lost by almost 9\% by the aggressive version 
of the same prefetcher. Most modern prefetchers are getting increasingly 
aggressive as that enables them to capture the complex memory access patterns 
seen in pointer-chasing applications. Thus, we need a mechanism that is 
orthogonal to the underlying prefetching scheme and can be used
to prune out the useful prefetcher from the useless ones.

Moreover, the on-line confidence mechanism used by most prefetchers is very
rudimentary. For example, SPP's internal confidence mechanism is based on taking 
the ratio C\textsubscript{d} = C\textsubscript{delta} / C\textsubscript{sig}.
This confidence was used to make the decision of whether to prefetch or not to
prefetch; and which level to prefetch.  While this approximation was shown to
work in the original implementation, we believe that a better form of
generalized on-line decision making was possible.  Hence, it was necessary to
build a robust and adaptable learning mechanism to accept / reject the
prefetch suggestions; and to decide the fill level (L2 Cache vs Last Level
Cache).

@@

Perceptrons in microarchitecture were first introduced by Jimenez and Lin~\cite{Perc_Branch} 
for branch prediction. In this work, a perceptron weights vector is selected by a hash of 
the branch address and its dot product with an input vector of the previous branch history 
outcomes. If the product is at least 0, the branch is predicted as taken, otherwise it is 
predicted not taken. The weights are updated with the perceptron update rule: if the prediction 
is incorrect, or the dot product does not exceed some threshold magnitude, then the weights 
are incremented or decremented based on their correlation to the corresponding history bits. 
Perceptron learning yields to superior accuracy, which results in increased opportunity for 
optimizations.

\begin{figure}[ht]
\includegraphics[width=\columnwidth]{Perc_Model}
\caption{The Basic Perceptron Model}
\label{Fig:Perc_Model}
\end{figure}

@@

In a perceptron-based mechanism, a uniform perceptron update principle is followed. As
shown in the figure~\ref{Fig:Perc_Model} Perceptrons 
are a simple model of neurons in neural networks \cite{perceptrons1,perceptrons2} modeled by 
vectors of signed weights learned through online training. The output of a perceptron is the 
dot product of the weights and a vector of inputs. In this work, we make use of the perceptron 
learning algorithm. There are two components to the abstract perceptron learning algorithm:

### 2018-12-05 conc.tex
Adding Perc image, motivation section

\section{Conclusion}
\label{Conclusion}
In this paper, we introduced perceptron-based learning for data prefetching.
The perceptron filter acts independently as a check on the underlying prefetch
engine.  PPF improves performance by over XX\% over the baseline,
corresponding to YY\% over the next best performing prefetcher.  We believe
that this is a robust and adaptable technique that can be used to enhance any
existing or upcoming prefetchers.

>>>>

\section{Conclusion}
\label{Conclusion}
In this paper, we introduced perceptron-based learning for data prefetching.
The perceptron filter acts independently as a check on the underlying prefetch
engine.  PPF improves performance of the base prefetcher used in isolation, 
by 11.9\%.  We believe that this is a robust and adaptable technique 
that can be used to enhance any existing and upcoming prefetchers.
